---
title: "dbplyr example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set-up connection and copy dataset

```{r}
library(dplyr)
con <- DBI::dbConnect(RSQLite::SQLite(), dbname = ":memory:")

# con <- DBI::dbConnect(odbc::odbc(), "Postgres_dev")

copy_to(con, nycflights13::flights, "flights",
        temporary = FALSE, 
        indexes = list(
          c("year", "month", "day"), 
          "carrier", 
          "tailnum",
          "dest"
        )
)
```

## Do some sample queries

```{r}
flights_db <- tbl(con, "flights")

flights_db %>% select(year:day, dep_delay, arr_delay)

flights_db %>% filter(dep_delay > 240)

flights_db %>% 
  group_by(dest) %>%
  summarise(delay = mean(dep_time))
```


Note the warning:

> Warning: Missing values are always removed in SQL.
> Use `mean(x, na.rm = TRUE)` to silence this warning
> This warning is displayed only once per session.

## Define analysis query and translate to SQL

```{r}
tailnum_delay_db <- flights_db %>% 
  group_by(tailnum) %>%
  summarise(
    delay = mean(arr_delay),
    n = n()
  ) %>% 
  arrange(desc(delay)) %>%
  filter(n > 100)

tailnum_delay_db %>% show_query()
```

Note the warning:

> Warning: ORDER BY is ignored in subqueries without LIMIT
> ℹ Do you need to move arrange() later in the pipeline or use window_order() instead?

## Do the query

```{r}
system.time(tailnum_delay <- tailnum_delay_db %>% collect())
tailnum_delay
```

## Comparison SQL

```{r}
system.time({
  res <- DBI::dbSendQuery(con, "SELECT *
                                FROM (SELECT `tailnum`, AVG(`arr_delay`) AS `delay`, COUNT(*) AS `n`
                                FROM `flights`
                                GROUP BY `tailnum`)
                                WHERE (`n` > 100.0)")
  tailnum_delay_sql <- DBI::dbFetch(res)
  DBI::dbClearResult(res)
})
head(tailnum_delay_sql)
```


## Comparison load data first

Check memory usage before executing this chunk!

```{r}
pryr::mem_used()
system.time({
  flights <- flights_db %>% collect()
  tailnum_delay_local <- flights %>% 
  group_by(tailnum) %>%
  summarise(
    delay = mean(arr_delay),
    n = n()
  ) %>% 
  arrange(desc(delay)) %>%
  filter(n > 100)
})
tailnum_delay_local
pryr::mem_used()
```

When processing locally, one first transfers about `r format(object.size(flights), units = "Mb")`
of data from the DB. Processing in the DB requires only about
`r format(object.size(tailnum_delay), units = "Kb")` of data transfer.



## Close connection

```{r}
DBI::dbDisconnect(con)
```